This bubble map, "Top Diseases by US City," visualizes the dominant diseases and their associated death counts across various US cities, with bubble size indicating death count (1×10\^5 to 4×10\^5) and color representing the disease.

### Key observations:

-   Measles (Green): Predominantly impacts northern cities with high death counts.
-   Chicago and New York City show very large bubbles, indicating substantial Measles-related deaths.
-   Philadelphia, Los Angeles, and Dallas also show Measles, but with lower death counts.
-   Pneumonia and Influenza (Blue): More prevalent in southern and western cities, generally with lower death counts.
-   Houston has the largest blue bubble, while cities like San Diego, Phoenix, San Antonio, and Jacksonville show smaller impacts.
-   Chlamydia (Red): Only noted in New York City with a small bubble, indicating a lower death count compared to other diseases shown. \### Loading and Visualizing Data

```{r Plotting}
#Add the city names for clarity
# Load required libraries
library(ggplot2)
library(maps)
library(readr)
library(ggrepel)

# Read the data
data <- read_csv("coordinated_cities.csv")

# Load US state map data
us_map <- map_data("state")

# Plot the map
ggplot() +
  # Draw the US map
  geom_polygon(data = us_map, aes(x = long, y = lat, group = group),
               fill = "gray90", color = "white") +
  # Plot disease bubbles (color = disease, size = count)
  geom_point(data = data, aes(x = Longitude, y = Latitude, 
                              size = `Death Count`, color = `Top Disease`),
             alpha = 0.8) +
  # Add city names near the bubbles
  geom_text_repel(data = data, 
                  aes(x = Longitude, y = Latitude, label = City),
                  size = 3.8, box.padding = 0.3, segment.color = "grey40") +
  # Scale bubble sizes
  scale_size_continuous(range = c(4, 20), name = "Death Count") +
  # Aesthetics
  labs(title = "Top Diseases by US City (Bubble Size = Death Count)",
       x = "Longitude", y = "Latitude", color = "Disease") +
  theme_minimal() +
  theme(legend.position = "right")

```

This choropleth map displays the total death counts by US state, using a color gradient from light to dark blue, where darker shades indicate higher death tolls (ranging from 250,000 to over 750,000).

### Key observations:

-   Highest Death Counts (Darkest Blue):
-   New York and California show the highest total deaths, exceeding 750,000.
-   High Death Counts (Darker Blue/Purple): -Texas, Florida, Pennsylvania, Illinois, Ohio, and Michigan exhibit significant death counts (between 500,000 and 750,000).
-   Moderate Death Counts (Medium Blue):
-   Numerous states across the Southeast, Midwest, Northwest, and Southwest fall into this category (250,000 to 500,000 deaths).
-   Lowest Death Counts (Lightest Blue):
-   Less populous states, primarily in the Rocky Mountains and Great Plains, show the lowest total deaths, closer to the 250,000 mark. \### Loading and Visualizing Data

```{r Plotting}
# Load required packages
library(ggplot2)
library(dplyr)
library(readr)
library(maps)

# Read the CSV file
state_data <- read_csv("state_death_totals.csv")  # Use your actual file path

# Prepare state map data
us_states <- map_data("state")

# Convert state names to lowercase to match map_data
state_data <- state_data %>%
  mutate(state_name = tolower(state_name))

# Join death data with map data
map_df <- left_join(us_states, state_data, by = c("region" = "state_name"))

# Plot choropleth map
ggplot(map_df, aes(long, lat, group = group, fill = total_deaths)) +
  geom_polygon(color = "white") +
  coord_fixed(1.3) +
  scale_fill_gradient(low = "lightblue", high = "darkblue", name = "Total Deaths") +
  labs(
    title = "US State-wise Total Deaths",
    x = "Longitude",
    y = "Latitude"
  ) +
  theme_minimal()
```

---
title: "Most_deaths"
output: html_document
---

### Key Observations

This bar chart, "Comparison of Total Deaths by Disease," ranks diseases by total deaths (in millions), categorized as Communicable (blue) or Non-Communicable (red).

-   Dominant Communicable Diseases: The chart highlights that communicable diseases account for the vast majority of deaths.
-   Tuberculosis is the leading cause (over 0.6 million deaths).
-   Pneumonia (approx. 0.42 million) and Pneumonia And Influenza (approx. 0.25 million) are also major contributors.
-   Other communicable diseases like Diphtheria, Typhoid Fever, and Influenza also show significant, though lower, death counts.
-   Minor Non-Communicable Impact: Pellagra is the sole non-communicable disease shown, with a negligible death count.
-   Low Impact Diseases: Diseases such as Chickenpox, Cholera, Yellow Fever, Dengue, and Leprosy show almost no deaths on this scale.

---
title: "Disease Deaths Comparison"
output: html_document
date: "2025-05-08"
---

## Analysis of Disease Deaths

This report compares total deaths across different diseases, marking them as communicable (C) or non-communicable (NC) using a bar chart.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)

# Print current working directory for debugging
cat("Current working directory:", getwd(), "\n")
```

### Loading and Visualizing Data

We load the data from the CSV file and create a bar chart to compare deaths by disease.

```{r data_plot, echo=TRUE}
# Attempt to read the CSV file with error handling
tryCatch({
  data <- read.csv("diseases_with_c_nc.csv")
  
  # Create a bar chart
  ggplot(data, aes(x = reorder(Disease, -Total_Deaths), y = Total_Deaths / 1000000, fill = Communicable)) +
    geom_bar(stat = "identity") +
    labs(title = "Comparison of Total Deaths by Disease",
         x = "Disease", y = "Total Deaths (in Millions)",
         fill = "Communicable (C/NC)") +
    scale_y_continuous(breaks = seq(0, max(data$Total_Deaths / 1000000), by = 0.1)) +  # Clear breaks in millions
    scale_fill_manual(values = c("C" = "skyblue", "NC" = "salmon")) +  # Color by communicability
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels for readability
}, error = function(e) {
  cat("Error: Could not read the file 'disease_deaths.csv'. Please ensure the file exists in the working directory (", getwd(), ") or provide the correct file path.\n")
  stop(e)
})
```

This lollipop chart, "Top 5 Disease By Deaths," visualizes the total deaths (in millions) for five diseases across all cities.

### Key Observations

-   Measles is the leading cause, with 2.0 million deaths.
-   Tuberculosis ranks second at approximately 1.2 million deaths.
-   Scarlet Fever is third, accounting for about 0.9 million deaths.
-   Pneumonia and Diphtheria follow, with approximately 0.7 million and 0.6 million deaths, respectively.

### Loading and Visualizing Data

```{r data_plot, echo=TRUE}

# Load required libraries
library(ggplot2)
library(dplyr)

# Read data from CSV
data <- read.csv("top_5_diseases.csv")

# Convert 'number' to millions for better readability on the plot
data <- data %>% mutate(number_in_millions = number / 1000000)

# Create a lollipop chart
ggplot(data, aes(x = reorder(disease, number), y = number_in_millions)) +
  geom_segment(aes(x = reorder(disease, number), xend = reorder(disease, number), 
                   y = 0, yend = number_in_millions), 
               color = "grey50", linewidth = 1.5) +  # Lines (sticks) of the lollipop
  geom_point(size = 6, aes(color = disease)) +  # Points (lollipops) with vivid colors
  scale_color_manual(values = c("Measles" = "#FF6F61", 
                                "Tuberculosis" = "#6B5B95", 
                                "Scarlet Fever" = "#88B04B", 
                                "Pneumonia" = "#F7CAC9", 
                                "Diphtheria" = "#92A8D1")) +  # Vivid color palette
  coord_flip() +  # Flip coordinates for better readability of disease names
  labs(title = "Top 5 Disease By Deaths",
       x = "Disease",
       y = "Total Deaths (in Millions)",
       caption = "Data represents total deaths across all cities.") +
  theme_minimal(base_size = 14) +
  theme(legend.position = "none",  # Remove legend since colors are for aesthetics
        plot.title = element_text(hjust = 0.5, face = "bold", color = "#333333"),
        axis.text = element_text(color = "#333333"),
        panel.grid.major.y = element_line(color = "grey80", linetype = "dashed"),
        panel.grid.minor = element_blank())  # Clean, lucid theme

# Save the plot (optional)
ggsave("disease_comparison_lollipop.png", width = 8, height = 6)
```
